The effects of varying depth in cosmic shear surveys

### 2019-10-10 paperdraft.tex
added the first batch of catherine's suggestion

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies\footnote{Instead of using the actual number of galaxies, we take the effective number density times respective area, as defined in \citet{Kuijken:2015}. Due to this we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.} per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via %\citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\cdot\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytic expression is derived in App. \ref{sec:model_e}.

@@

@@

\section{Results}
\label{sec:results}
Here we apply both the analytic and the semi-analytic method to model the data from the KiDS+VIKING-450 survey \citep[KV450,][]{Wright:2018}. While the application of the semi-analytic method is straightforward, for the analytic method we need to decide how to estimate the weight function $W$ from the given redshift data. Following the separation of a survey into quantiles as in Sect. \ref{sec:xipm_semianalytic}, we define $W(\b\theta)\equiv W_n$ whenever $\b\theta$ is in a pointing of quantile $n$. For the determination of $W_n$ we test two approaches:
 As a first method, following \citet{2006APh....26...91V,1997A&A...322....1B}, we estimate 
\begin{equation}
W_n \propto \la z \ra _n^{0.85}\, ,
\end{equation}
where $\la z\ra_n$ is the average redshift of quantile $n$. As a second method we define \begin{equation}
W_n \propto \sqrt{ \la\gamma_{\rm{t}}\gamma_{\rm{t}}\ra (\theta_{\rm{ref}}) + \la \gamma_\times\gamma_\times\ra (\theta_{\rm{ref}})} = \sqrt{\xi_{+,nn}^{ij}(\theta_{\rm{ref}})}\, ,
\end{equation}
where the $\xi_{+,nn}^{ij}(\theta_{\rm{ref}})$ denotes the model correlation function defined in Sect. \ref{sec:xipm_semianalytic}, evaluated at a characteristic scale $\theta_{\rm{ref}}$, that needs to be chosen.

@@

We applied both analytic methods and the semianalytic one to data from KV450, following the separation into tomographic bins as in H18, and computed the ratio between our model for the observed correlation functions, $\xi_{\pm}^{\rm{obs}}$, and the model correlation functions, $\xi_{\pm}$, ignoring variations in depth.

We compare our models to mock KiDS-like data created using a modified version of the Full-sky Lognormal Astro-fields Simulation Kit \citep[\sc{FLASK},][Joachimi, Lin, et al. in prep.]{Xavier:2016}.  Using {\sc{FLASK}} lognormal fields, we generate galaxy mocks with coherent clustering and lensing signals.   Adopting a linear relation between the limiting $r$-band magnitude and the effective number density, fit to the KiDS data, we imprint the variable depth of the full KiDS-1000 footprint \citep{Kuijken:2019} on a {\sc{Healpix}}\footnote{\url{https://healpix.sourceforge.io}} grid of Nside=4096 (add cite for Healpix).  Our resolution choice represents a compromise between minimising the mock computation time and maximising the accuracy of the recovered shear signal at small angular scales.  With Nside=4096 $\xi_+$ is accurate to 7\% above ~1 arcmin, and $\xi_-$ is accurate to 10\% above ~6 arcmin.   For each \sc{Healpix} pixel in the KiDS-1000 data, the limiting $r$-band magnitude defines the effective number density of sources, and the average source redshift distribution for each tomographic redshift bin (see Fig.~\ref{fig:nofz_quantiles}).  In the uniform depth case, redshifts and number densities are sampled from the average of these tomographic sets.  The ratio of the lensing signals from the two mocks is computed, averaged over 2000 shape-noise-free realizations, and is shown in Fig.~\ref{fig:all_xis} (red).

>>>>

Let $N^i(\b \theta),N^j(\b\theta)$ be the average weighted number of galaxies\footnote{Instead of using the actual number of galaxies, we take the effective number density, as defined in \citet{Kuijken:2015}, scaled by the respective survey area. Due to this we account for different weighting of galaxies in the shear correlation functions as well as in the average redshift distribution.} per pointing in redshift bins $i$ and $j$ and let $W^i(\b \theta),W^j(\b\theta)$ be the weighting of average shear. The observed correlation functions $\xi^{ij,\text{obs}}_\pm(\theta)$ now change from uniform depth, $\xi_\pm^{ij,\rm{uni}}(\theta)$, via %\citep[see][]{2002A&A...396....1S}
\begin{align}
\xi^{ij,\text{obs}}_\pm(\theta) = & \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_{\rm t}(\bth)\gamma^{j,\rm{obs}}_{\rm t}(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \\
 & \pm \frac{\la N^i(\bth)N^j(\bthp)\gamma^{i,\rm{obs}}_\times(\bth)\gamma^{j,\rm{obs}}_\times(\bthp)\ra }{\la N^i(\bth)N^j(\bthp)\ra} \nonumber\\
 = & \frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i(\bth)N^j(\bthp)\ra} \xi_{\pm}^{ij,\rm{uni}}(\theta) \, ,
 \label{eq:xipmblub1}
 \end{align}
 where the average $\la\cdot\ra$ represents both an ensemble average as well as an average over the position $\bth$.
 Assuming that the depth of different pointings is uncorrelated, the only important property of a galaxy pair is whether or not they lie in the same pointing. We want to denote the probability that a random galaxy pair of separation $\theta$ lies in the same pointing by $E(\theta)$. This function is depicted in Fig. \ref{fig:eoftheta_lin}, and an analytic expression is derived in App. \ref{sec:model_e}.

@@

\label{eq:xipm_analytic}

@@

\section{Results}
\label{sec:results}
We compare the analytic and semi-analytic models for a variable-depth cosmic shear measurement in a KiDS-like survey. While the application of the semi-analytic method is straightforward, for the analytic method we need to decide how to estimate the weight function $W$ from the given redshift data. Following the separation of a survey into quantiles as in Sect. \ref{sec:xipm_semianalytic}, we define $W(\b\theta)\equiv W_n$ whenever $\b\theta$ is in a pointing of quantile $n$. For the determination of $W_n$ we test two approaches:
 As a first method, following \citet{2006APh....26...91V,1997A&A...322....1B}, we estimate 
\begin{equation}
W_n \propto \la z \ra _n^{0.85}\, ,
\end{equation}
where $\la z\ra_n$ is the average redshift of quantile $n$. As a second method we define \begin{equation}
W_n \propto \sqrt{ \la\gamma_{\rm{t}}\gamma_{\rm{t}}\ra (\theta_{\rm{ref}}) + \la \gamma_\times\gamma_\times\ra (\theta_{\rm{ref}})} = \sqrt{\xi_{+,nn}^{ij}(\theta_{\rm{ref}})}\, ,
\end{equation}
where the $\xi_{+,nn}^{ij}(\theta_{\rm{ref}})$ denotes the model correlation function defined in Sect. \ref{sec:xipm_semianalytic}, evaluated at a characteristic scale $\theta_{\rm{ref}}$, that needs to be chosen.

@@

In this Section we calculate both the analytic (Eq.~\ref{eq:xipm_analytic}) and semi-analytic (Eq.~\ref{eq:correctionfunction2}) models for the shear correlation function from a KiDS-like variable depth survey.   We adopt the tomographic bins defined in H18 and their resulting best-fit cosmological parameters to present, in Fig.~\ref{fig:all_xis}, the ratio between our models for the observed correlation functions $\xi_\pm^{\rm{obs}}$, and the standard theoretical prediction that assumes uniform depth. We find that the level of variation in the depth of the KiDS survey increases the amplitude of the observed shear correlation function, on sub-pointing scales, by up to 5\% relative to the uniform-depth case.

We compare our models to mock KiDS-like data created using a modified version of the Full-sky Lognormal Astro-fields Simulation Kit \citep[\sc{FLASK},][Joachimi, Lin, et al. in prep.]{Xavier:2016}.  Using {\sc{FLASK}} lognormal fields, we generate galaxy mocks with coherent clustering and lensing signals.   Adopting a linear relation between the limiting $r$-band magnitude and the effective number density, fit to the KiDS data, we imprint the variable depth of the full KiDS-1000 footprint \citep{Kuijken:2019} on a {\sc{Healpix}}\footnote{\url{https://healpix.sourceforge.io}} grid of Nside=4096 (add cite for Healpix).  Our resolution choice represents a compromise between minimising the mock computation time and maximising the accuracy of the recovered shear signal at small angular scales.  With Nside=4096 $\xi_+$ is accurate to 7\% above ~1 arcmin, and $\xi_-$ is accurate to 10\% above ~6 arcmin.   For each {\sc{Healpix}} pixel in the KiDS-1000 data, the limiting $r$-band magnitude defines the effective number density of sources, and the average source redshift distribution for each tomographic redshift bin (see Fig.~\ref{fig:nofz_quantiles}).  In the uniform depth case, redshifts and number densities are sampled from the average of these tomographic sets.  The ratio of the lensing signals from the two mocks is computed, averaged over 2000 shape-noise-free realizations, and is shown in Fig.~\ref{fig:all_xis} (red).
